Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)

### 2022-04-11 doc/manuscript/ESSDmanuscript.Rmd
misc changes

#Statistics
n_records_ForC= NA # total n records in ForC
n_independent_records_ForC= NA #suspected duplicates removed
n_records_EFDB_variables= NA #n independent records relevant to EFDB
n_records_sent_to_EFDB = NA # n records sent to EFDB
n_records_posted_EFDB = NA # n records accepted and posted in EFDB
```

@@

The International Panel on Climate Change (IPCC) provides guidance for national greenhouse gas inventories for reporting to the United Nations Framework Convention on Climate Change [UNFCCC, @ipcc_2019_2019].
Under this guidance, greenhouse gas inventories include all managed land, including most of the world's forest land [@ogle_delineating_2018].
The IPCC inventory guidelines include specific instructions for accounting for greenhouse gas (mainly CO~2~) exchanges between forest land and the atmosphere [@ipcc_agriculture_2006; @ipcc_2019_2019].
This guidance has improved over the years as more of the relevant underlying data has become available [@requenasuarez_estimating_2019; @rozendaal_aboveground_2022], but there remains room for continuous improvement as the science advances.
For example, the year following the release of the latest IPCC guidelines, @cook-patton_mapping_2020 found that the latest default rates may underestimate rates of C accumulation in regrowth forests by 32% on average and fail to capture eight-fold variation within ecozones. 
In addition, @cuni-sanchez_high_2021 found that aboveground C stocks in mature African tropical montane forests were two-thirds higher than the IPCC default values for these forests.
This rapid evolution of scientific information on the climate mitigation potential of forests is beneficial to climate mitigation efforts, but requires improved mechanisms for communicating the latest information from scientific researchers to the practitioners who need reliable estimates for greenhouse gas mitigation planning. 
Moreover, high variability of forest C cycling within ecozones [e.g., @cook-patton_mapping_2020; @refs] implies that it is useful for practitioners to have access to locally-specific information, when available.

@@

The Global Forest Carbon Database, ForC ([https://forc-db.github.io/](https://forc-db.github.io/)), is the largest collection of published estimates of forest carbon stocks, increments, and annual fluxes [@anderson-teixeira_forc_2018; @anderson-teixeira_carbon_2021].
For C currently contains **##** records from **##** plots in **##** distinct geographical areas, along with records of stand age and disturbance history.
As such, ForC is positioned to improve forest C accounting through the transfer of data to EFDB.
The purpose of this publication is to document that process and provide recommendations for future improvements.

@@

For quantifying forest role in global C cycle, we ultimately care about:
(1) C stocks -- stores of C that would be vulnerable to release to the atmosphere upon land use change
(2) C increments -- changes in those C stocks.

## Carbon stocks
Forest ecosystem C stocks may be parsed into pools in various ways. 
IPCC parses into biomass (aboveground and belowground), dead organic matter (dead wood and litter), and soil organic matter (Table `r table_pools`).
Quantifying these requires a one-time measurement.

>>>>

#Statistics
n_records_ForC= NA # total n records in ForC
n_plots_ForC = NA # total n plots in ForC
n_areas_ForC = NA # total n geographic areas in ForC
n_independent_records_ForC= NA #suspected duplicates removed
n_records_EFDB_variables= NA #n independent records relevant to EFDB
n_records_sent_to_EFDB = NA # n records sent to EFDB
n_records_posted_EFDB = NA # n records accepted and posted in EFDB
```

@@

The International Panel on Climate Change (IPCC) provides guidance for national greenhouse gas inventories for reporting to the United Nations Framework Convention on Climate Change [UNFCCC, @ipcc_2006_2006;@ipcc_2019_2019].
Under this guidance, greenhouse gas inventories include all managed land, including most of the world's forest land [@ogle_delineating_2018].
The IPCC inventory guidelines include specific instructions for accounting for greenhouse gas (mainly CO~2~) exchanges between forest land and the atmosphere [@ipcc_agriculture_2006; @ipcc_2019_2019].
This guidance has improved over the years as more of the relevant underlying data has become available [@requenasuarez_estimating_2019; @rozendaal_aboveground_2022], but there remains room for continuous improvement as the science advances.
For example, the year following the release of the latest IPCC guidelines, @cook-patton_mapping_2020 found that the latest default rates may underestimate rates of C accumulation in regrowth forests by 32% on average and fail to capture eight-fold variation within ecozones. 
In addition, @cuni-sanchez_high_2021 found that aboveground C stocks in mature African tropical montane forests were two-thirds higher than the IPCC default values for these forests.
This rapid evolution of scientific information on the climate mitigation potential of forests is beneficial to climate mitigation efforts, but requires improved mechanisms for communicating the latest information from scientific researchers to the practitioners who need reliable estimates for greenhouse gas mitigation planning. 
Moreover, high variability of forest C cycling within ecozones [e.g., @cook-patton_mapping_2020; @refs] implies that it is useful for practitioners to have access to locally-specific information, when available.

@@

The Global Forest Carbon Database, ForC ([https://forc-db.github.io/](https://forc-db.github.io/)), is the largest collection of published estimates of forest carbon stocks, increments, and annual fluxes [@anderson-teixeira_forc_2018; @anderson-teixeira_carbon_2021].
For C currently contains `r n_records_ForC` records from `r n_plots_ForC` plots in `r n_areas_ForC` distinct geographical areas, along with records of stand age and disturbance history.
As such, ForC is positioned to improve forest C accounting through the transfer of data to EFDB.
The purpose of this publication is to document that process and provide recommendations for future improvements.

@@

The end goal of IPCC greenhouse gas inventories is to quantify greenhouse gas emissions to, or withdrawals from, the atmosphere on an annual basis, most commonly for the purpose of national inventories [@ipcc_2006_2006;@ipcc_2019_2019]. 
More proximately, IPCC guidelines focus on:
(1) C stocks -- i.e., stores of C that would be vulnerable to release to the atmosphere upon land use change
and (2) C increments -- i.e., changes in those C stocks over one or more years.
*[detail how these are used]*

## Carbon stocks
Forest ecosystem C stocks may be parsed into pools in various ways, and while certain definitions and thresholds are more common than others, there is no single standard for measuring or reporting that is adhered to by all -- or even most -- studies. 
IPCC parses into biomass (aboveground and belowground), dead organic matter (dead wood and litter), and soil organic matter (Table `r table_pools`).
Quantifying these requires a one-time measurement.

### 2022-04-11 doc/manuscript/ESSDmanuscript.tex
misc changes

The International Panel on Climate Change (IPCC) provides guidance for
national greenhouse gas inventories for reporting to the United Nations
Framework Convention on Climate Change
\citep[UNFCCC,][]{ipcc_2019_2019}. Under this guidance, greenhouse gas
inventories include all managed land, including most of the world's
forest land \citep{ogle_delineating_2018}. The IPCC inventory guidelines
include specific instructions for accounting for greenhouse gas (mainly
CO\textsubscript{2}) exchanges between forest land and the atmosphere
\citep{ipcc_agriculture_2006, ipcc_2019_2019}. This guidance has
improved over the years as more of the relevant underlying data has
become available
\citep{requenasuarez_estimating_2019, rozendaal_aboveground_2022}, but
there remains room for continuous improvement as the science advances.
For example, the year following the release of the latest IPCC
guidelines, \citet{cook-patton_mapping_2020} found that the latest
default rates may underestimate rates of C accumulation in regrowth
forests by 32\% on average and fail to capture eight-fold variation
within ecozones. In addition, \citet{cuni-sanchez_high_2021} found that
aboveground C stocks in mature African tropical montane forests were
two-thirds higher than the IPCC default values for these forests. This
rapid evolution of scientific information on the climate mitigation
potential of forests is beneficial to climate mitigation efforts, but
requires improved mechanisms for communicating the latest information
from scientific researchers to the practitioners who need reliable
estimates for greenhouse gas mitigation planning. Moreover, high
variability of forest C cycling within ecozones
\citep[e.g.,][]{cook-patton_mapping_2020, refs} implies that it is
useful for practitioners to have access to locally-specific information,
when available.

@@

The Global Forest Carbon Database, ForC
(\url{https://forc-db.github.io/}), is the largest collection of
published estimates of forest carbon stocks, increments, and annual
fluxes
\citep{anderson-teixeira_forc_2018, anderson-teixeira_carbon_2021}. For
C currently contains \textbf{\#\#} records from \textbf{\#\#} plots in
\textbf{\#\#} distinct geographical areas, along with records of stand
age and disturbance history. As such, ForC is positioned to improve
forest C accounting through the transfer of data to EFDB. The purpose of
this publication is to document that process and provide recommendations
for future improvements.

@@

For quantifying forest role in global C cycle, we ultimately care about:
(1) C stocks -- stores of C that would be vulnerable to release to the
atmosphere upon land use change (2) C increments -- changes in those C
stocks.

@@

Forest ecosystem C stocks may be parsed into pools in various ways. IPCC
parses into biomass (aboveground and belowground), dead organic matter
(dead wood and litter), and soil organic matter (Table 1). Quantifying
these requires a one-time measurement.

>>>>

The International Panel on Climate Change (IPCC) provides guidance for
national greenhouse gas inventories for reporting to the United Nations
Framework Convention on Climate Change
\citep[UNFCCC,][]{ipcc_2006_2006, ipcc_2019_2019}. Under this guidance,
greenhouse gas inventories include all managed land, including most of
the world's forest land \citep{ogle_delineating_2018}. The IPCC
inventory guidelines include specific instructions for accounting for
greenhouse gas (mainly CO\textsubscript{2}) exchanges between forest
land and the atmosphere \citep{ipcc_agriculture_2006, ipcc_2019_2019}.
This guidance has improved over the years as more of the relevant
underlying data has become available
\citep{requenasuarez_estimating_2019, rozendaal_aboveground_2022}, but
there remains room for continuous improvement as the science advances.
For example, the year following the release of the latest IPCC
guidelines, \citet{cook-patton_mapping_2020} found that the latest
default rates may underestimate rates of C accumulation in regrowth
forests by 32\% on average and fail to capture eight-fold variation
within ecozones. In addition, \citet{cuni-sanchez_high_2021} found that
aboveground C stocks in mature African tropical montane forests were
two-thirds higher than the IPCC default values for these forests. This
rapid evolution of scientific information on the climate mitigation
potential of forests is beneficial to climate mitigation efforts, but
requires improved mechanisms for communicating the latest information
from scientific researchers to the practitioners who need reliable
estimates for greenhouse gas mitigation planning. Moreover, high
variability of forest C cycling within ecozones
\citep[e.g.,][]{cook-patton_mapping_2020, refs} implies that it is
useful for practitioners to have access to locally-specific information,
when available.

@@

The Global Forest Carbon Database, ForC
(\url{https://forc-db.github.io/}), is the largest collection of
published estimates of forest carbon stocks, increments, and annual
fluxes
\citep{anderson-teixeira_forc_2018, anderson-teixeira_carbon_2021}. For
C currently contains NA records from NA plots in NA distinct
geographical areas, along with records of stand age and disturbance
history. As such, ForC is positioned to improve forest C accounting
through the transfer of data to EFDB. The purpose of this publication is
to document that process and provide recommendations for future
improvements.

@@

The end goal of IPCC greenhouse gas inventories is to quantify
greenhouse gas emissions to, or withdrawals from, the atmosphere on an
annual basis, most commonly for the purpose of national inventories
\citep{ipcc_2006_2006, ipcc_2019_2019}. More proximately, IPCC
guidelines focus on: (1) C stocks -- i.e., stores of C that would be
vulnerable to release to the atmosphere upon land use change and (2) C
increments -- i.e., changes in those C stocks over one or more years.
\emph{{[}detail how these are used{]}}

@@

Forest ecosystem C stocks may be parsed into pools in various ways, and
while certain definitions and thresholds are more common than others,
there is no single standard for measuring or reporting that is adhered
to by all -- or even most -- studies. IPCC parses into biomass
(aboveground and belowground), dead organic matter (dead wood and
litter), and soil organic matter (Table 1). Quantifying these requires a
one-time measurement.
